Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm

### 2022-04-13 final_metrics_analysis.Rmd

from aif360.algorithms.preprocessing import DisparateImpactRemover
from BlackBoxAuditing.repairers.GeneralRepairer import Repairer

@@

  cols = fair_metrics.columns.values
  obj = fair_metrics.loc['objective']
  size_rect = [0.2,0.2,0.2,0.4,0.25]
  rect = [-0.1,-0.1,-0.1,0.8,0]
  bottom = [-1,-1,-1,0,0]
  top = [1,1,1,2,1]
  bound = [[-0.1,0.1],[-0.1,0.1],[-0.1,0.1],[0.8,1.25]]

@@

@@

@@

# diremover_rp = diremover.fit_transform(X_test)
# AttributeError: 'DataFrame' object has no attribute 'features' (probem with .fit_transform function)
# .fit_transform() angry because of X_test...

```

@@

```{python post-processing generalized entropy error}
gen_entr_error_CEO = generalized_entropy_error(y_test, y_pred_CEO, alpha=1, pos_label = 0)

print(f'[Calibrated Equalized Odds] The generalized entropy error is {gen_entr_error_CEO: .2}')
```

```{python post-processing fairness metrics plot}
plot_fair_metrics([stat_par_diff_CEO, eq_opp_diff_CEO, gen_entr_error_CEO], 'Calibrated Equalized Odds')
plt.show()
```

>>>>

# from aif360.algorithms.preprocessing import DisparateImpactRemover
# from BlackBoxAuditing.repairers.GeneralRepairer import Repairer

@@

  cols = fair_metrics.columns.values
  obj = fair_metrics.loc['objective']
  size_rect = [0.2,0.2,0.2,0.4]
  rect = [-0.1,-0.1,-0.1,0.8]
  bottom = [-1,-1,-1,0]
  top = [1,1,1,2]
  bound = [[-0.1,0.1],[-0.1,0.1],[-0.1,0.1],[0.8,1.25]]

@@

### Plot Fairness Metrics (Baseline)

@@

### Plot Fairness Metrics (Reweighting)

@@

# diremover_rp = diremover.fit_transform(X_test)
# AttributeError: 'DataFrame' object has no attribute 'features' (probem with .fit_transform function)
# .fit_transform() angry because of X_test... 
```

@@

```{python post-processing average odds difference}
avg_odds_diff_CEO = average_odds_difference(y_test, y_pred_CEO, prot_attr='race', pos_label = 0)

print(f'[Calibrated Equalized Odds] The average odds difference is {avg_odds_diff_CEO: .2}')
```

```{python post_processing disparate impact ratio}
disp_impact_ratio_CEO = disparate_impact_ratio(y_test, y_pred_CEO, prot_attr='race', priv_group = 1, pos_label = 0)

print(f'[Calibrated Equalized Odds] The disparate impact ratio is {disp_impact_ratio_CEO: .2}')
```

### Plot Fairness Metrics (Calibrated Equalized Odds)

```{python post-processing fairness metrics plot}
plot_fair_metrics([stat_par_diff_CEO, eq_opp_diff_CEO, avg_odds_diff_CEO, disp_impact_ratio_CEO], 'Calibrated Equalized Odds')
plt.show()
```
